### 2015-01-27 paper.tex
more stuff

Constraint grammar was first introduced by \cite{KarlssonTODO}. 
It a tool for disambiguating output by morphological analyser.

>>>>

Constraint Grammar (CG) was first introduced by \cite{KarlssonTODO}. 
It a tool for disambiguating output by morphological analyser.

Parsing is an action where an unstructured input is transformed into a
structured output. In a sense, applying CG rules on a text shouldn't
be considered as parsing: it doesn't really build a new structure,
such as phrase structure tree or dependency tree.
The grammatical information is already in the input, and applying the
rules leaves (ideally) only the correct options. 
It doesn't require that the text should be parsed as full units; it is
neither top-down or bottom up. Rules can have a context of specified
amount of words (e.g. 1 to the left; 2 to the right), unlimited amount
of words (e.g. anywhere to the left; 2 to the right and further from
that) or they can be limited by a barrier (e.g. anywhere to the right
until we see an adverb).

In the example below, we show possible tags for the sentence ``the bear sleeps''.

\begin{lstlisting}
<the>     Det
<bear>    Noun, Sg
          Verb, Pl, Pres
          Verb, Inf
<sleep>   Noun, Pl
          Verb, Sg3, Pres
\end{lstlisting}

We could disambiguate this sentence with two rules: 
\begin{enumerate}
\item Select noun (or remove verb) after determiner;
\item Select verb, if is preceded by noun and followed by a sentence boundary.
\end{enumerate}

We would go through the input twice. The first rule would fire at the
first pass: we find a case where there is something tagged as
\texttt{Det} followed by something with \texttt{Noun} among the
analyses. We will apply the rule where we select the noun reading, and
discard the two other options (present tense verb or nonfinite verb).

Then, at the next pass, we have two unambiguous tokens,
and we go through the input again, to see
if the first pass has created chances for new rules to
activate. Now we find a context, where there is an ambiguous word
preceded by a noun (\emph{bear}) and followed by a sentence boundary,
so we can apply rule 2. The finished analysis is as follows:

\begin{lstlisting}
<the>     Det
<bear>    Noun, Sg
<sleep>   Verb, Sg3, Pres
\end{lstlisting}

CG can also be applied to shallow syntax. After having disambiguated
the text morphologically, we can apply possible syntactic labels. For
instance, a noun can have a syntactic role of subject, object or
predicative; and a verb in present tense 3rd person singular can be
only the main verb (possibly combined with a conjunction, e.g. ``the
bear eats and sleeps''). The next step is shown below; the analyses
chosen in the previous step are shown in the angle brackets, and
syntactic labels are fully capitalised and preceded by @.

\begin{lstlisting}
<the,  Det>                
<bear, N, Sg>           @SUBJ
                        @OBJ
                        @PREDC
<sleep, V, Sg3, Pres>   @MAINV
\end{lstlisting}
